The \project{unRAVE} catalog

### 2016-09-05 article/unrave.tex
Include text about GES comparison

There are 142 stars that overlap between \rave\ and the fourth internal
data release of the \ges\ survey. These are a mix of main-sequence, 
sub-giant and red giant branch stars.  Despite most of these stars having
relatively low S/N ratios in \rave\ ($\approx 25$), there is good 
agreement in stellar parameters (Figure~\ref{fig:ges-comparison}.  
The RMS in effective temperature, surface gravity and metallicity is
236~K, 0.39~dex, and 0.18~dex, respectively. 
% TODO: Check if those numbers above are still valid when final catalog is used.

\stub{A paragraph about the abundance agreement.}

@@

\acknowledgements
We thank 
	Jonathan Bird (Vanderbilt),
	Sven Buder (MPIA), 
and 
	Sergey Koposov (Cambridge).
This research made use of: 
  	NASA's Astrophysics Data System Bibliographic Services;
  	Astropy, a community-developed core Python package for Astronomy \citep{astropy};
and 
  	\project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for K.~H.~ has been provided through the Simons 
Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.

>>>>

There are 142 stars that overlap between \rave\ and the fourth internal
data release of the \ges\ survey. These are a mix of main-sequence, 
sub-giant and red giant branch stars.  Despite most of these stars having
relatively low S/N ratios in \rave\ ($\approx 25$), there is good 
agreement in stellar parameters (Figure~\ref{fig:ges-stellar-parameters}.  
The RMS in effective temperature, surface gravity and metallicity is
233~K, 0.37~dex, and 0.17~dex, respectively.

Based on this comparison, we find no evidence for a systematic offset 
between stars on the main-sequence and those on the giant branch. Specifically,
we do not find any dwarf stars that are misclassified as sub-giants, as we
found for some metal-poor stars in Section~\ref{sec:validation-gold-standards}.
In addition we find no evidence for a systematic offset in metallicity (or
individual abundances) between giant stars and main-sequence stars.  This is
a crucial inference, as the metallicities for stars in our main-sequence
training set have a principally different source to those on the giant branch.
We cannot make these same inferences based on other surveys, like \apogee,
because (1) \apogee\ stars formed part of the training set, and (2) they do
not include main-sequence stars.  Even if we found good agreement between
\epic\ and \apogee\ metallicities, this would not be informative, because
\apogee\ is the source of metallicity for many stars on the giant branch in
the \epic\ sample. Therefore, although this is a qualitative comparison only,
it is reassuring that there is no obvious systematic difference between the
metallicities of main-sequence and giant branch stars.

The metallicity agreement between this work and \ges\ extends down to low
metallicity, near $[{\rm Fe/H}] \approx -1.5$.  The scatter increases for
the four stars in the overlap sample with [Fe/H] $< -1$, however these
spectra have low S/N ratios, which is reflected in the larger uncertainties
reported for these metallicities.
% TODO: Check if that is still the case after we examine errors from repeat
% 		obs.

The fourth internal data release of the \ges\ includes detailed chemical
abundances of up to 45~species ($\approx32$ elements at different ionization
stages).  This provides us with an independent validation for our detailed
abundances on the giant branch.  These comparisons are shown for six elemental
abundances in Figure~\ref{fig:ges-abundances}, where markers are colored by the
S/N of the \rave\ spectrum.  The number of stars available in each abundance
comparison varies due to what is available in the \ges\ data release, which
is itself a function of the instrument used, the spectral type, and other
factors.  The absolute bias for individual elements varies from as low as
0.01~dex ([Al/H]) to as high as 0.22~dex ([Si/H]), where we over-estimate 
[Si/H] abundances relative to the \ges\ survey. The RMS deviation in each
label is small for stars with [X/H] $\sim -0.5$, before increasing at lower
metallicities.  If we consider all stars, the smallest abundance RMS we see
with respect to \ges\ is 0.18~dex for [Ca/H].  The increasing RMS at low 
metallicity is likely a consequence of multiple factors, namely: inaccurate
abundance labels for metal-poor stars (Section~\ref{sec:the-training-set});
only weak, blended lines being available in \rave, which cease to be visible
in hot and/or metal-poor stars; and to a lesser extent, low S/N ratios for 
those particular stars being compared.  Unfortunately not all of these
factors are represented by the quoted errors in each label.  For these reasons, 
although it affects only a small number of stars, we recommend caution when 
using individual abundances for very metal-poor giant stars in our sample.

@@

\acknowledgements
We thank 
	Jonathan Bird (Vanderbilt),
	Sven Buder (MPIA), 
	Sergey Koposov (Cambridge),
and 
	Sofia Randich (INAF).
This research made use of: 
  	NASA's Astrophysics Data System Bibliographic Services;
  	Astropy, a community-developed core Python package for Astronomy \citep{astropy};
and 
  	\project{TOPCAT} \citep{Taylor_2005}.
This work was partly supported by the European Union FP7 programme through ERC 
grant number 320360. Funding for K.~H.~ has been provided through the Simons 
Foundation Society of Fellows and the Marshall Scholarship.
Funding for RAVE has been provided by: the Australian Astronomical Observatory; 
the Leibniz-Institut fuer Astrophysik Potsdam (AIP); the Australian National 
University; the Australian Research Council; the French National Research Agency;
the German Research Foundation (SPP 1177 and SFB 881); the European Research 
Council (ERC-StG 240271 Galactica); the Istituto Nazionale di Astrofisica at 
Padova; The Johns Hopkins University; the National Science Foundation of the USA
(AST-0908326); the W. M. Keck foundation; the Macquarie University; the 
Netherlands Research School for Astronomy; the Natural Sciences and Engineering 
Research Council of Canada; the Slovenian Research Agency; the Swiss National 
Science Foundation; the Science \& Technology Facilities Council of the UK; 
Opticon; Strasbourg Observatory; and the Universities of Groningen, Heidelberg 
and Sydney. The RAVE web site is https://www.rave-survey.org.
